Internal Gravity Wave Breaking in White Dwarf Binaries

### 2019-11-07 writeups/Su_IGW_break.tex
okay actual draft for Daniel

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for a range of astrophysical problems. They are
the most important sources of gravitational waves (GWs) for the Laser
Interferometric Space Antenna (LISA) \citep{lisa}. They are also thought to
produce interesting optical transients such as underluminous supernovae
\citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal novae
\citep{tidal_novae}. Most importantly, they have been proposed as the likely
progenitors of type Ia supernovae (e.g. \citealt{Ia0,webbink} or more recently
\citealt{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
expand the catalog to a few hundreds \citep{lsst_wd} \citep[results based on
\emph{Gaia}'s second data release have already begun to
appear][]{gaiaDD,gaiaDD2}, and the Large Synoptic Survey Telescope (LSST, first
light scheduled for 2020) will likely detect a few thousand more
\citep{lsst_wd}. These observations will significantly advance the understanding
of WD binaries and their evolution.

@@

\citealt{drazin, klostermeyer, winters1994} describe the onset of wave breaking.
At intermediate amplitudes, wave breaking occurs via triadic resonances which
transfer energy from the ``parent’’ internal gravity wave (in our problem, the
tidally forced wave) to ``daughter’’ internal gravity waves on smaller length
scales that efficiently damp. The horizontal momentum flux decreases from $F$ to
$0$ over this breaking region. The lost flux is deposited into mean flow
\begin{equation}
    \overline{U}(z, t) \equiv \ev*{u_x}_x.\label{eq:mean_flow}
\end{equation}
As the mean flow grows, a \emph{critical layer} may form, discussed below.

@@

% TODO stash here
\begin{equation}
    \overline{U}_c \equiv \omega/k_x,
\end{equation}
The further evolution of the critical layer with the incident IGW is
thought to be responsible for tidal synchronization in stellar binaries
\citep{zahn75,gn89} as well as terrestrial phenomena such as the quasi-biennial
oscillation \citep{lindzen_qbo}.

The behavior of an IGW incident upon a critical layer was first studied in the
inviscid, linear regime in \citep{booker_bretherton}, which found nearly
complete absorption of the IGW\@. In particular, the incident wave has amplitude
reflection and transmission coefficients that can be expressed in terms of the
local Richardson number Ri at the height of the critical layer $z_c$:
\begin{align}
    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\overline{U}}{z}}^2}}_{z_c},
        \label{eq:ri_def}\\
    \mathcal{R}_A &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}.
        \label{eq:crit_coeffs}
\end{align}
In the $\mathrm{Ri} \gg 1$ limit, $\mathcal{R}, \mathcal{T} \ll 1$
and the incident wave is completely absorbed to a good approximation.

This result extends to viscous fluids \citep{hazel}, but weakly nonlinear theory
\citep{brown_stewartson} and numerical simulations \citep{winters1994} suggest
that nonlinear effects may significantly enhance reflection and transmission.

@@

We aim to damp waves that reach the edge of the simulation domain without
inducing nonphysical reflection. To do so, we replace material derivatives
in \autoref{se:nl_upsilon} with:
\begin{align}
    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\nonumber\\
    \Gamma(z) &= \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
        + \tanh \frac{z_B - z}{\Delta z}},\label{eq:Gamma}
\end{align}
where $z_B = 1.5H, z_T = 10.5H$ are the boundaries of the damping zone.
This strongly damps perturbations below $z_B$ and above $z_T$ with damping time
$\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
governed by $\Delta z$. We use $\Delta z = 0.25H$ and $\tau = 1/15N$. This
prescription is similar to \citep{lecoanet_damp} and has the advantage of being
smooth, important for spectral methods. Further details of our implementation of
the fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.

@@

For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we expect
$\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones, i.e.\
$z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are defined in
\autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For consistency
with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t = z_b + H$
(using $z_t = z_T - \Delta z$ just below the damping layer instead does not
change the results). The resulting measurement of $\hat{A}_i(t)$ is shown in
\autoref{fig:lin_amps}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/lin_amps.png}
    \caption{Amplitude of excited wave over time in weak forcing simulation,
    computed using \autoref{eq:ahat_def}. $\hat{A}_i(t) = 1$ corresponds to
    perfect agreement with the analytical estimate. After an initial transient
    phase, we observe great agreement with \autoref{eq:uz_lin}, and in
    particular $\hat{A}_i(t)$ asymptotes to $\approx 1$, implying continuous
    excitation of identical IGW with the expected
    amplitude.}\label{fig:lin_amps}
\end{figure}

@@

\section{Wave Breaking Simulations}\label{s:sim}

@@

In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $F / F_{al}$ as a function
of $z$ at the times depicted in \autoref{fig:snapshots}. While the behavior of
$\overline{U}$ seems to conform qualitatively with the predictions of
\autoref{ss:crit_layer}, the behavior of $F / F_{al}$ exhibits two notable
features: (i) the incident flux seems to fluctuate greatly with time, and (ii)
there seems to be a small transmitted feature at many of the later times. We
will discuss further these features in \autoref{ss:reflectivity}, after first
analyzing the propagation of the critical layer.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
    \caption{Plot of $\overline{U}(z, t)$ and $F(z, t) / F_{al}(z)$ respectively
    at the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
    1024$ simulation. $\overline{U}, F$ follow their definitions in
    \autoref{eq:mean_flow} and \autoref{eq:F_def}. $\overline{U}$ is given in
    units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
    flow velocity (see \autoref{ss:crit_layer}). The propagation of the critical
    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
    are evident.}\label{fig:nl_fluxes}
\end{figure}

@@

For the simulation shown in \autoref{fig:nl_fluxes}, we may analyze the location
of the critical layer. As $F$ fluctuates significantly in the simulation, we
measure the location of the critical layer using an average of where flux
deposition occurs
\begin{align}
    z_{c, \min} &= \argmin_z \z{z: F(z) > 0.3F_{al}},\nonumber\\
    z_{c, \max} &= \argmax_z \z{z: F(z) < 0.3F_{al}},\nonumber\\
    z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
\end{align}
This was found to a relatively stable estimator of the critical layer location.
Using other estimators does not significantly change the results of the
analysis.

The evolution of $z_c$ depends on the deposited flux $\Delta F(t)$ per
\autoref{eq:zc_anal}. To measure $\Delta F(t)$ across the critical layer $z_c$
from the data, we defined:
\begin{align}
    F_{<}(t) &= \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
        ,\label{eq:sbelow_def}\\
    F_>(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
        \label{eq:sabove_def}\\
    \Delta F(t) &\equiv F_<(t) - F_{>}(t).\label{eq:ds_def}
\end{align}
$\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
Below the critical layer, we average over an interval of length
$\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
necessary to make a measurement of the incident flux unaffected by the
turbulence within the critical layer itself. The height of the critical layer is
limited by $\mathrm{Ri} \lesssim 1$, which bounds its vertical extent $\sim
\frac{1}{k_{z}}$. We empirically found an offset of $\Delta z = \frac{3}{k_z}$
was necessary to be sufficiently far from strong fluctuations near the critical
layer.

Above the critical layer, waves decay over a much smaller vertical extent as
they have larger $z$ wavenumbers thanks to the shear flow Doppler shift.
Accurate measurement of these fluctuations above the critical layer is important
to measuring the transmission feature accurately, seen in
\autoref{fig:nl_fluxes} to be weak and attenuate quickly. We average only where
$F < 0$ since the decay scale of the trasmitted flux varies in time (see
\autoref{fig:nl_fluxes} again), so this was found to be an adequate way of
measuring only the transmitted flux.

Finally, we may plot the measured $z_c$ against two semi-analytic predictors:
(i) integration of \autoref{eq:zc_anal} using the measured $\Delta F(t)$, and
(ii) substituting the time-averaged $\ev{\Delta F}_t$ (over the entire length of
the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both stable and less
well-defined at early times (when the critical layer is thick and transient
behavior is still strong), we instead integrate backwards from the end of the
simulation, using $z_c(t_f)$ as initial condition. The resulting predictors are
depicted in \autoref{fig:nl_front}. The good agreement between the evolution of
$z_c(t)$ and its estimate via $\Delta F(t)$ and \autoref{eq:zc_anal} are
noteworthy.

Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
0.71F_{al}$ predictor, which serves mostly as a fiducial in the plot. The
general agreement clearly demonstrates $\Delta F < F_{al}$ and that momentum
flux absorption is incomplete. However, its underprediction of critical layer
velocity at early times contrasts with its overprediction at later times. This
suggests that momentum flux absorption changes significantly over time, which is
in line with the observations in the subsequent \autoref{ss:reflectivity}.

@@

Since fluid instabilities are local, $\mathrm{Ri}$ must be measured in the
immediate vicinity of a fluid parcel. Thus, we first assign an $\mathrm{Ri}$ for
every $x$ in the critical layer, then take the median as $\mathrm{Ri}$ for the
entire layer. To avoid noisiness, the local $\mathrm{Ri}$ is computed using the
vertical distance over which the local $u_x$ increases from $0.3 \times$ its
critical value to its critical value. The value $0.3$ effectively excludes the
self-acceleration induced by the IGW\@. This can be written
\begin{align}
    z_{CL, \min}(x, t) &= \argmin_\zeta
        \z{z: u_x(x, \zeta, t) > 0.3\overline{U}_c},\nonumber\\
    z_{CL, \max}(x, t) &= \argmax
        \zeta \z{z: u_x(x, \zeta, t) < \overline{U}_c},\nonumber\\
    \mathrm{Ri}(t) &\equiv
        \med_x\p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
            \overline{U}_c)^2}}.\label{ss:ri_med_def}
\end{align}
To understand the variation in $\mathrm{Ri}$ over $x$, we can also plot using
the minimum over $x$ (the maximum is significantly noisier). Both of these are
shown in \autoref{fig:nl_f_ri}. The quick evolution of $\mathrm{Ri}$ to its
saturated value is reflective of the fact that IGW are \emph{anti-diffusive}.
This property is a simple consequence of the IGW dispersion relation
\autoref{eq:disp_rel}, which is approximately $\omega k_z \approx Nk_x$ such
that as an IGW propagates into a shear flow, $\omega$ decreases and $k_z$
increases, enhancing dissipation and further driving mean flow acceleration.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_f_ri.png}
    \caption{Local Richardson number of the flow at the critical layer over time
    as defined in \autoref{ss:ri_med_def}. The dotted red and solid green lines
    denote respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the
    maximum greatly exceeds the vertical scale of this plot). These numbers
    effectively measure the mean and spread in width of the critical layer over
    $x$. Note that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so
    this plot suggests the shear at the critical layer does not steepen past the
    KHI onset.}\label{fig:nl_f_ri}
\end{figure}

@@

Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
winding is used to enforce solid body rotation on the grounds that $t_A \gg
t_{gw}$, where
\begin{equation}
    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}}.
\end{equation}
the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV})
measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
inspiral timescale. Before solid body rotation is attained, another relevant
timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
synchronization timescale is
\begin{align}
    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\nonumber\\
        &\sim 2 \times 10^5\;\mathrm{yr}
            \p{\frac{M}{M_{\odot}}}
            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
            \p{\frac{R}{R_{\oplus}}}^2
            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
\end{align}
A representative $\tau$ has been taken from \citep{bukart}. However, since only
the outer $\sim 10^{-4}M_{\odot}$ need be heated for the thermodynamically
interesting effects studied in \citep{fullerIV} and \citep{tidal_novae}, it
seems that magnetic winding cannot absolutely rule out energetic outbursts such
as tidal novae resulting from strong shear flows.

@@

    \item Solar-type stars (inner conbmtive, outer radiative): different
        equation of state/stratification but could be qualitatively similar.

>>>>

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for a range of astrophysical problems. They are
the most important sources of gravitational waves (GWs) for the Laser
Interferometric Space Antenna (LISA) \citep{lisa}. They are also thought to
produce interesting optical transients such as underluminous supernovae
\citep{underlum}, Ca-rich fast transients \citep{carich}, and tidal novae
\citep{tidal_novae}. Most importantly, they have been proposed as the likely
progenitors of type Ia supernovae (e.g. \citealp{Ia0,webbink} or more recently
\citealp{Ia1,Ia2}). While presently only a few tens of compact WD binaries are
known \citep{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
expand the catalog to a few hundreds \citep{lsst_wd} \citep[results based on
\emph{Gaia}'s second data release have already begun to
appear][]{gaiaDD,gaiaDD2}, and the Large Synoptic Survey Telescope (LSST, first
light scheduled for 2020) will likely detect a few thousand more
\citep{lsst_wd}. These observations will significantly advance the understanding
of WD binaries and their evolution.

@@

\citealp{drazin, klostermeyer, winters1994} describe the onset of wave breaking.
At intermediate amplitudes, wave breaking occurs via triadic resonances which
transfer energy from the ``parent’’ internal gravity wave (in our problem, the
tidally forced wave) to ``daughter’’ internal gravity waves on smaller length
scales that efficiently damp. The horizontal momentum flux decreases from $F$ to
$0$ over this breaking region. The lost flux is deposited into a horizontal mean
flow
\begin{equation}
    \overline{U}(z, t) \equiv \ev{u_x}_x.\label{eq:mean_flow}
\end{equation}
As the mean flow grows, a \emph{critical layer} may form, discussed below.

@@

A horizontal shear flow $\overline{U}(z, t)\hat{x}$ enters the fluid equations
via the Lagrangian derivative, which can be decomposed as
\begin{equation}
    \md{}{t} = \pd{}{t} + \overline{U} \pd{}{x} + \p{\bm{u}' \cdot \vec{\nabla}},
\end{equation}
where $\bm{u}'$ is the velocity field \emph{without} the shear flow, for
clarity. Thus, $\overline{U}$ has the effect of Doppler shifting the time
derivative into the frame comoving with the mean flow. If $\overline{U}$ is
roughly constant, then the behavior of a small plane wave perturbation satisfies
modified dispersion relation
\begin{equation}
    \p{\omega - \overline{U}k_x}^2 =
        \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
        \label{eq:disp_rel_U}
\end{equation}
This is just \autoref{eq:disp_rel} with $\omega \to \omega - \overline{U}k_x$.

It is apparent that if $\overline{U} = \overline{U}_c$, where
\begin{equation}
    \overline{U}_c \equiv \frac{\omega}{k_x},
\end{equation}
then the dispersion relation is singular and the linear solution breaks down.
Physically, this corresponds to the Doppler-shifted frequency of the IGW being
zero. Anywhere $\overline{U} = \overline{U}_c$ is called a \emph{critical
layer}.

The behavior of an IGW incident upon a critical layer was first studied in the
inviscid, linear regime in \citealp{booker_bretherton}, which found nearly
complete absorption of the IGW\@. In particular, the incident wave has amplitude
reflection and transmission coefficients that can be expressed in terms of the
local Richardson number Ri at the height of the critical layer $z_c$:
\begin{align}
    \mathrm{Ri} &\equiv \at{\frac{N^2}{\p{\pd{\overline{U}}{z}}^2}}_{z_c},
        \label{eq:ri_def}\\
    \mathcal{R}_A &= e^{-2\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}, &
    \mathcal{T}_A &= e^{-\pi \sqrt{\mathrm{Ri} - \frac{1}{4}}}.
        \label{eq:crit_coeffs}
\end{align}
In the $\mathrm{Ri} \gg 1$ limit, $\mathcal{R}, \mathcal{T} \ll 1$
and the incident wave is completely absorbed to a good approximation.

This result extends to viscous fluids \citep{hazel}, but weakly nonlinear theory
\citep{brown_stewartson} and numerical simulations \citep{winters1994} suggest
that nonlinear effects may significantly enhance reflection and transmission.
Understanding the interaction of IGW and critical layers has important
applications: it is thought to be responsible for tidal synchronization in
stellar binaries \citep{zahn75,gn89} as well as terrestrial phenomena such as
the quasi-biennial oscillation \citep{lindzen_qbo}.

@@

We aim to damp waves that reach the edge of the simulation domain without
inducing nonphysical reflection. To do so, we replace material derivatives
in \autoref{se:nl_upsilon} with:
\begin{align}
    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\\
    \Gamma(z) &= \frac{1}{2\tau}\s{2 + \tanh \frac{z - z_T}{\Delta z}
        + \tanh \frac{z_B - z}{\Delta z}},\label{eq:Gamma}
\end{align}
where $z_B = 1.5H, z_T = 10.5H$ are the boundaries of the damping zone.
This strongly damps perturbations below $z_B$ and above $z_T$ with damping time
$\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
governed by $\Delta z$. We use $\Delta z = 0.25H$ and $\tau = 1/15N$. This
prescription is similar to \citep{lecoanet_damp} and has the advantage of being
smooth, important for spectral methods. Further details of our implementation of
the fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.

@@

For a simulation satisfying $\xi_z k_z \ll 1$ everywhere, we expect
$\hat{A}_i(t) = 1$ when integrated between the forcing and damping zones, i.e.\
$z_b \gtrsim z_0, z_t \lesssim z_T$ ($z_0, z_T$ are defined in
\autoref{eq:vol_drive} and \autoref{eq:Gamma} respectively). For consistency
with the nonlinear case later, we choose $z_b = z_0 + 3\sigma, z_t = z_b + H$
(using $z_t = z_T - \Delta z$ just below the damping layer instead does not
change the results). The resulting measurement of $\hat{A}_i(t)$ is shown in
\autoref{fig:lin_amps}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/lin_amps.png}
    \caption{Amplitude of excited wave over time in weak forcing simulation,
    computed using \autoref{eq:ahat_def}. $\hat{A}_i(t) = 1$ corresponds to
    perfect agreement with the analytical estimate. After an initial transient
    phase, we observe great agreement with \autoref{eq:uz_lin}, and in
    particular $\hat{A}_i(t)$ asymptotes to $\approx 1$, implying continuous
    excitation of identical IGW with the expected amplitude. The small
    suppression may be due to timestepping errors, as a relatively large fixed
    step size $\Delta t = 0.1/N$ was used for this
    simulation.}\label{fig:lin_amps}
\end{figure}

@@

\section{Numerical Simulations of Wave Breaking}\label{s:sim}

@@

In \autoref{fig:nl_fluxes}, we plot $\overline{U}$ and $F / F_{al}$ as a function
of $z$ at the times depicted in \autoref{fig:snapshots}.
At each time, $\overline{U}$ is close to zero below the critical layer, but then
sharply increases to $\overline{U}_c$ at the critical layer. Above the critical
layer, $\overline{U}$ varies slightly due to momentum transport within the
spun-up layer. This is in agreement with the predictions
of\autoref{ss:crit_layer}.

Similarly, $F \lesssim F_{al}$ below the critical layer, and then decreases to
about zero above the critical layer. However, two notable deviations can be
observed: (i) the incident flux on the critical layer seems to fluctuate
somewhat temporally, and (ii) there seems to be a small transmitted feature at
many of the later times. We will discuss further these features in
\autoref{ss:reflectivity}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
    \caption{Plot of $\overline{U}(z, t)$ and $F(z, t) / F_{al}(z)$ respectively
    at the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
    1024$ simulation. $\overline{U}, F$ follow their definitions in
    \autoref{eq:mean_flow} and \autoref{eq:F_def}. $\overline{U}$ is given in
    units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
    flow velocity (see \autoref{ss:crit_layer}). The propagation of the critical
    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
    are evident.}\label{fig:nl_fluxes}
\end{figure}

@@

For our fiducial simulation, we may analyze the location of the critical layer
and compare to \autoref{eq:zc_anal}. As $F$ fluctuates significantly in the
simulation, we measure the location of the critical layer using an average of
where flux deposition occurs
\begin{align}
    z_{c, \min} &\equiv \argmin_z \z{z: F(z) > 0.3F_{al}},\\
    z_{c, \max} &\equiv \argmax_z \z{z: F(z) < 0.3F_{al}},\\
    z_c &\equiv \frac{z_{c, \min} + z_{c, \max}}{2}.\label{eq:zc_def}
\end{align}
This was found to a relatively stable estimator of the critical layer location.
Using other estimators does not significantly change the results of the
analysis.

We also need to measure $\Delta F(t)$ across the critical layer $z_c$ from the
data, so we define
\begin{align}
    F_{<}(t) &\equiv \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
        ,\label{eq:sbelow_def}\\
    F_>(t) &\equiv \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
        \label{eq:sabove_def}\\
    \Delta F(t) &\equiv F_<(t) - F_{>}(t).\label{eq:ds_def}
\end{align}
$\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
Below the critical layer, we average over an interval of length
$\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
necessary to make a measurement of the incident flux unaffected by the
turbulence within the critical layer itself. The height of the critical layer is
limited by $\mathrm{Ri} \lesssim 1$ (see \autoref{ss:khi}), which bounds its
vertical extent $\sim \frac{1}{k_{z}}$. We empirically found an offset of
$\Delta z = \frac{3}{k_z}$ was necessary to be sufficiently far from strong
fluctuations near the critical layer.

Above the critical layer, waves decay over a much smaller vertical extent as
they have larger $z$ wavenumbers thanks to the shear flow Doppler shift.
Accurate measurement of these fluctuations above the critical layer is important
to measuring the transmission feature accurately, seen in
\autoref{fig:nl_fluxes} to be weak and attenuate quickly. We average only where
$F < 0$ since the decay scale of the trasmitted flux varies in time (see
\autoref{fig:nl_fluxes} again), so this was found to be an adequate way of
measuring the transmitted flux before it dissipates.

Finally, we may plot the measured $z_c$ against two semi-analytic predictors:
(i) integration of \autoref{eq:zc_anal} using the measured $\Delta F(t)$, and
(ii) substituting the time-averaged $\ev{\Delta F}_t$ (over the entire length of
the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both unstable and
less well-defined at early times (when the critical layer is thick and transient
behavior is still strong), we instead integrate backwards from the end of the
simulation, using $z_c(t_f)$ as initial condition. The resulting predictors are
depicted in \autoref{fig:nl_front}. The good agreement between the evolution of
$z_c(t)$ and its estimate via $\Delta F(t)$ and \autoref{eq:zc_anal} are
noteworthy.

Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
0.71F_{al}$ predictor, which serves mostly as a fiducial in the plot. The
general agreement clearly demonstrates $\Delta F < F_{al}$ and that momentum
flux absorption is incomplete. However, its underprediction of critical layer
velocity at early times contrasts with its overprediction at later times. This
suggests that momentum flux absorption changes significantly over time, in
agreement with \autoref{ss:reflectivity}.

@@

Since fluid instabilities are local, $\mathrm{Ri}$ must be measured in the
immediate vicinity of a fluid parcel. Thus, we first assign an $\mathrm{Ri}$ for
every $x$ in the critical layer, then take the median as $\mathrm{Ri}$ for the
entire layer. To avoid noisiness, the local $\mathrm{Ri}$ is computed using the
vertical distance over which the local $u_x$ increases from $0.3 \times$ its
critical value to its critical value. The value $0.3$ is necessary to exclude
the small mean flow generated in the weakly nonlinear regime well below the
critical layer. This is written
\begin{align}
    z_{CL, \min}(x, t) &\equiv \argmin_z
        \z{z\mid u_x(x, z, t) > 0.3\overline{U}_c},\\
    z_{CL, \max}(x, t) &\equiv \argmax_z
        \z{z\mid u_x(x, z, t) < \overline{U}_c},\\
    \mathrm{Ri}(t) &\equiv
        \med_x\p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
            \overline{U}_c)^2}}.\label{eq:ri_med_def}
\end{align}
To understand the variation in $\mathrm{Ri}$ over $x$, we can also plot using
the minimum over $x$ (the maximum is significantly noisier). Both of these are
shown in \autoref{fig:nl_f_ri}. The quick evolution of $\mathrm{Ri}$ to its
saturated value is reflective of the fact that IGW are \emph{anti-diffusive}.
This property is a simple consequence of the IGW dispersion relation
\autoref{eq:disp_rel}, which is approximately $\omega k_z \approx Nk_x$ such
that as an IGW propagates into a shear flow, $\omega$ decreases and $k_z$
increases, enhancing dissipation and further driving mean flow acceleration.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_f_ri.png}
    \caption{Local Richardson number of the flow at the critical layer over time
    as defined in \autoref{eq:ri_med_def}. The dotted red and solid green lines
    denote respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the
    maximum greatly exceeds the vertical scale of this plot). These numbers
    effectively measure the mean and spread in width of the critical layer over
    $x$. Note that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so
    this plot suggests the shear at the critical layer does not steepen past the
    KHI onset.}\label{fig:nl_f_ri}
\end{figure}

@@

Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
winding is used to enforce solid body rotation on the grounds that $t_A \gg
t_{gw}$, where
\begin{equation}
    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
        \sim 10^2\;\mathrm{yr}\p{\frac{10^3\;\mathrm{G}}{B}},
\end{equation}
the Alfv\'en wave crossing time (evaluated for a CO WD in \citealp{fullerIV}),
measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
inspiral timescale. Before solid body rotation is attained, another relevant
timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
tidal forcing frequency $\sigma = m\p{\Omega - \Omega_{spin}}$, we note that
angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
synchronization timescale is
\begin{align}
    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\\
        &\sim 2 \times 10^5\;\mathrm{yr}
            \p{\frac{M}{M_{\odot}}}
            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
            \p{\frac{R}{R_{\oplus}}}^2
            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
\end{align}
A representative $\tau$ has been taken from \citep{bukart}. However, since only
the outer $\sim 10^{-4}M_{\odot}$ need be heated for the thermodynamically
interesting effects studied in \citep{fullerIV} and \citep{tidal_novae}, it
seems that magnetic winding cannot absolutely rule out energetic outbursts such
as tidal novae resulting from strong shear flows.

@@

    \item Solar-type stars (inner convective, outer radiative): different
        equation of state/stratification but could be qualitatively similar.
